---
title: "Longitudinal STIs"
author: "David R. Pletta, MPH"
date: '2022-07-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Formatting output for knitr
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)

#Setting the working directory for the markdown document
knitr::opts_knit$set(root.dir = "/Users/davidpletta/Dropbox/ R Projects/LEGACY/DFRR PCORI/AIM I - Longitudinal STI/Longitudinal STIs")
```

```{r, echo = FALSE}
library(tidyverse)
library(tidylog)
library(haven)
```

## Importing and Cleaning Data

```{r}
#Importing EMR cohorts from R dataset
load("EMR Cohorts_LEGACY_07202022.RData")
```

```{r}
general_cohort$Age[general_cohort$Age > 100] <- NA
```

```{r}
table(general_cohort$Age)
```

```{r}
#Need to recode STI positive variable to include anogenital STIs (not any STI)
#general_cohort <- subset(general_cohort, select = -c(STI_Positive))

general_cohort$STI_Pos_Anogenital <- ifelse(is.na(general_cohort$Last_Pos_CT_Cervical) == FALSE |
                             is.na(general_cohort$Last_Pos_CT_Urethral) == FALSE |
                             is.na(general_cohort$Last_Pos_CT_Rectal) == FALSE |
                             is.na(general_cohort$Last_Pos_GC_Cervical) == FALSE |
                             is.na(general_cohort$Last_Pos_GC_Urethral) == FALSE |
                             is.na(general_cohort$Last_Pos_GC_Rectal) == FALSE, 1, 0)

general_cohort$STI_Pos_Anogenital <- factor(general_cohort$STI_Pos_Anogenital)

levels(general_cohort$STI_Pos_Anogenital) <- c("No", "Yes")

table(general_cohort$STI_Pos_Anogenital, general_cohort$Year)
```

```{r}
general_cohort_2016 <- general_cohort[which(general_cohort$Year == 2016),]
```

```{r}
general_cohort_2016$STI_Pos_Urethral <- ifelse((is.na(general_cohort_2016$Last_Pos_CT_Urethral) == FALSE) | (is.na(general_cohort_2016$Last_Pos_GC_Urethral) == FALSE), 1, 0)

table(general_cohort_2016$STI_Pos_Urethral)
```

```{r}
general_cohort_2016$STI_Pos_Rectal <- ifelse((is.na(general_cohort_2016$Last_Pos_CT_Rectal) == FALSE) | (is.na(general_cohort_2016$Last_Pos_GC_Rectal) == FALSE), 1, 0)

table(general_cohort_2016$STI_Pos_Rectal)
```

```{r}
general_cohort_2016$STI_Pos_Cervical <- ifelse((is.na(general_cohort_2016$Last_Pos_CT_Cervical) == FALSE) | (is.na(general_cohort_2016$Last_Pos_GC_Cervical) == FALSE), 1, 0)

table(general_cohort_2016$STI_Pos_Cervical)
```

```{r}
#Coding the number of anogenital STI tests in the emr dataset
general_cohort$STI_Test_Anogenital <- ifelse(is.na(general_cohort$Last_CT_Cervical) == FALSE |
                             is.na(general_cohort$Last_CT_Urethral) == FALSE |
                             is.na(general_cohort$Last_CT_Rectal) == FALSE |
                             is.na(general_cohort$Last_GC_Cervical) == FALSE |
                             is.na(general_cohort$Last_GC_Urethral) == FALSE |
                             is.na(general_cohort$Last_GC_Rectal) == FALSE, 1, 0)

general_cohort$STI_Test_Anogenital <- factor(general_cohort$STI_Test_Anogenital)

levels(general_cohort$STI_Test_Anogenital) <- c("No", "Yes")
```

```{r}
#Creating a categorical variable for age
general_cohort <- mutate(general_cohort, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

general_cohort$AgeCat <- factor(general_cohort$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))

```

```{r}
general_cohort$On_PrEP <- factor(general_cohort$On_PrEP)

levels(general_cohort$On_PrEP) <- c("No", "Yes")

table(general_cohort$On_PrEP, general_cohort$Year)
```

```{r}
general_cohort <- mutate(general_cohort, Gender = case_when(
  GenderID == "Female" ~ 0,
  GenderID == "Male" ~ 1,
  (GenderID == "Nonbinary") & (Sex_at_Birth == "Female") ~ 2,
  (GenderID == "Nonbinary") & (Sex_at_Birth == "Male") ~ 3
))

general_cohort$Gender <- factor(general_cohort$Gender)

levels(general_cohort$Gender) <- c("Female", "Male", "Nonbinary AFAB", "Nonbinary AMAB")

table(general_cohort$Gender)

prop.table(table(general_cohort$Gender))
```

```{r}
general_cohort <- mutate(general_cohort, RaceID = case_when(
  Race == "White" ~ 0,
  Race == "Black or African American" ~ 1,
  Race == "Multiracial" ~ 2,
  Race == "Another Race" | Race == "Asian" | Race == "American Indian or Alaska Native" | 
    Race == "Native Hawaiian or Other Pacific Islander" ~ 3
))

general_cohort$RaceID <- factor(general_cohort$RaceID)

levels(general_cohort$RaceID) <- c("White", "Black or African American", "Multiracial", "Another Race")

table(general_cohort$RaceID)
```

```{r}
general_cohort <- mutate(general_cohort, HIV_Risk = case_when(
  (HIV_Status == 0) & (On_PrEP == "No") ~ 0,
  (HIV_Status == 0) & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == 1) & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == 1) & (HIV_Suppressed == "200+ copies") ~ 3
))

general_cohort$HIV_Risk <- factor(general_cohort$HIV_Risk)

levels(general_cohort$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
table(general_cohort$HIV_Risk)

prop.table(table(general_cohort$HIV_Risk))
```

```{r}
general_cohort$HIV_Status <- factor(general_cohort$HIV_Status)

levels(general_cohort$HIV_Status) <- c("HIV-", "HIV+")

table(general_cohort$HIV_Status)
```

## Cohort Sociodemographics at Baseline (2016)

```{r}
#Subsetting baseline data for the general cohort
general_baseline <- general_cohort[which(general_cohort$Year == 2016),]
```

```{r}
median(general_baseline$Age, na.rm = TRUE)
IQR(general_baseline$Age, na.rm = TRUE)
```

```{r}
library(tableone)

general_baseline_table <- CreateTableOne(vars = c("Site", "Age", "AgeCat", "Sex_at_Birth", 
                                                  "Gender", "Race", "Ethnicity", 
                                                  "Insurance", "FPL_Level", "On_Hormones", 
                                                  "On_PrEP", "HIV_Status", "HIV_Suppressed", 
                                                  "HIV_Risk", "STI_Pos_Anogenital"),
                                         data = general_baseline, includeNA = TRUE)

general_baseline_table
```

## Hormone Descriptives at Baseline

```{r}
general_baseline$On_Blocker <- factor(general_baseline$On_Blocker)
levels(general_baseline$On_Blocker) <- c("No", "Yes")

general_baseline$On_Antiandrogen <- factor(general_baseline$On_Antiandrogen)
levels(general_baseline$On_Antiandrogen) <- c("No", "Yes")

general_baseline$On_Estrogen <- factor(general_baseline$On_Estrogen)
levels(general_baseline$On_Estrogen) <- c("No", "Yes")

general_baseline$On_Progesterone <- factor(general_baseline$On_Progesterone)
levels(general_baseline$On_Progesterone) <- c("No", "Yes")

general_baseline$On_Testosterone <- factor(general_baseline$On_Testosterone)
levels(general_baseline$On_Testosterone) <- c("No", "Yes")
```

```{r}
library(tableone)

general_baseline_hormones <- CreateTableOne(vars = c("On_Blocker", "On_Antiandrogen", "On_Estrogen", "On_Progesterone", "On_Testosterone"),
                                         data = general_baseline, includeNA = TRUE)

general_baseline_hormones
```

## Assessing Trends in the General Cohort

```{r}
#Examining attrition based on hormone status
table(general_cohort$On_Hormones, general_cohort$Year)

prop.table(table(general_cohort$On_Hormones, general_cohort$Year))
```

## Incidence of Positive Anogenital STI Tests

```{r}
table(general_cohort$STI_Test_Anogenital)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$STI_Pos_Anogenital)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$Year)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$Year)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$Site)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$Site)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$On_Hormones)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$On_Hormones)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$AgeCat, useNA = "always")
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$AgeCat, useNA = "always")
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$Gender)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$Gender)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$Race)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$Race)
```

```{r}
table(general_cohort$STI_Test_Anogenital, general_cohort$Ethnicity)
```

```{r}
table(general_cohort$STI_Pos_Anogenital, general_cohort$Ethnicity)
```

## Creating Complete-Case Dataset

```{r}
#Selecting variables to include in our complete-case dataset
general_cohort_completecase <- subset(general_cohort, select = c(ID, Year, Site, Age, Gender, Race, Ethnicity, Insurance, FPL_Level,
                                                                 On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, STI_Pos_Anogenital))

#Creating a complete-case dataset with patients who have STI testing data
general_cohort_completecase <- general_cohort_completecase[complete.cases(general_cohort_completecase),]
```

```{r}
general_cohort_completecase$ID <- as.integer(general_cohort_completecase$ID)
```

```{r}
general_cohort_completecase <- mutate(general_cohort_completecase, STI_Positive = case_when(
  STI_Pos_Anogenital == "No" ~ 0,
  STI_Pos_Anogenital == "Yes" ~ 1
))

#Creating a categorical variable for age
general_cohort_completecase <- mutate(general_cohort_completecase, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

general_cohort_completecase$AgeCat <- factor(general_cohort_completecase$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))


general_cohort_completecase <- mutate(general_cohort_completecase, RaceID = case_when(
  Race == "White" ~ 0,
  Race == "Black or African American" ~ 1,
  Race == "Multiracial" ~ 2,
  Race == "Another Race" | Race == "Asian" | Race == "American Indian or Alaska Native" | 
    Race == "Native Hawaiian or Other Pacific Islander" ~ 3
))

general_cohort_completecase$RaceID <- factor(general_cohort_completecase$RaceID)

levels(general_cohort_completecase$RaceID) <- c("White", "Black or African American", "Multiracial", "Another Race")

general_cohort_completecase <- mutate(general_cohort_completecase, Obs_Year = case_when(
  Year == 2016 ~ 0,
  Year == 2017 ~ 1,
  Year == 2018 ~ 2,
  Year == 2019 ~ 3
))

general_cohort_completecase <- mutate(general_cohort_completecase, HIV_Risk = case_when(
  (HIV_Status == "HIV-") & (On_PrEP == "No") ~ 0,
  (HIV_Status == "HIV-") & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "200+ copies") ~ 3
))

general_cohort_completecase$HIV_Risk <- factor(general_cohort_completecase$HIV_Risk)

levels(general_cohort_completecase$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")

```

## Complete-Case Models

```{r}
library(geepack)

CC_STI_unadjusted <- geeglm(STI_Positive ~ On_Hormones + Site + Obs_Year, family = poisson(link = "log"), data = general_cohort_completecase,
                            id = ID, waves = Obs_Year, corstr = "exchangeable")

summary(CC_STI_unadjusted)
```

```{r}
CC_STI_unadjusted_df <- df.residual(CC_STI_unadjusted)
```

```{r}
CC_STI_adjusted <- geeglm(STI_Positive ~ On_Hormones + Site + AgeCat + Gender + RaceID + Ethnicity + Insurance + FPL_Level + HIV_Risk + Obs_Year, 
                          family = poisson(link = "log"), data = general_cohort_completecase, id = ID, waves = Obs_Year, corstr = "exchangeable")

summary(CC_STI_adjusted)
```

```{r}
CC_STI_adjusted_df <- df.residual(CC_STI_adjusted)
```

## Creating Wide Dataset for Imputation

```{r}
#Creating separate datasets for baseline sociodemographics & time-varying demographics and outcomes
library(sqldf)

Cohort_Demographics <- sqldf('SELECT ID, Site, Age, Gender, 
                             Race, Ethnicity FROM general_cohort WHERE Year = 2016 ORDER BY ID')

Outcomes_2016 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, STI_Pos_Anogenital
FROM general_cohort WHERE Year = 2016 ORDER BY ID')

Outcomes_2017 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, STI_Pos_Anogenital
FROM general_cohort WHERE Year = 2017 ORDER BY ID')

Outcomes_2018 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, STI_Pos_Anogenital
FROM general_cohort WHERE Year = 2018 ORDER BY ID')

Outcomes_2019 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, STI_Pos_Anogenital
FROM general_cohort WHERE Year = 2019 ORDER BY ID')
```

```{r, message = FALSE, echo = FALSE}
#Renaming variables in the outcome datasets before proceeding to left joins.
library(data.table)

setnames(Outcomes_2016, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'STI_Pos_Anogenital'), 
         new = c('Insurance_2016', 'FPL_Level_2016', 'On_Hormones_2016', 'On_PrEP_2016', 
                                'HIV_Status_2016', 'HIV_Suppressed_2016', 'STI_Pos_Anogenital_2016'))

setnames(Outcomes_2017, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'STI_Pos_Anogenital'), 
         new = c('Insurance_2017', 'FPL_Level_2017', 'On_Hormones_2017', 'On_PrEP_2017', 
                                'HIV_Status_2017', 'HIV_Suppressed_2017', 'STI_Pos_Anogenital_2017'))

setnames(Outcomes_2018, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'STI_Pos_Anogenital'), 
         new = c('Insurance_2018', 'FPL_Level_2018', 'On_Hormones_2018', 'On_PrEP_2018', 
                                'HIV_Status_2018', 'HIV_Suppressed_2018', 'STI_Pos_Anogenital_2018'))

setnames(Outcomes_2019, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'STI_Pos_Anogenital'), 
         new = c('Insurance_2019', 'FPL_Level_2019', 'On_Hormones_2019', 'On_PrEP_2019', 
                                'HIV_Status_2019', 'HIV_Suppressed_2019', 'STI_Pos_Anogenital_2019'))                                
```

```{r}
#Left joining the datasets on patient ID

Demo_16 <- merge(Cohort_Demographics, Outcomes_2016, by = "ID")

Demo_1617 <- merge(Demo_16, Outcomes_2017, by = "ID", all.x = TRUE)

Demo_161718 <- merge(Demo_1617, Outcomes_2018, by = "ID", all.x = TRUE)

general_cohort_wide <- merge(Demo_161718, Outcomes_2019, by = "ID", all.x = TRUE)
```

```{r}
#Converting character variables to factor variables
general_cohort_wide$Gender <- factor(general_cohort_wide$Gender)
general_cohort_wide$Race <- factor(general_cohort_wide$Race)
general_cohort_wide$Ethnicity <- factor(general_cohort_wide$Ethnicity)

general_cohort_wide$Insurance_2016 <- factor(general_cohort_wide$Insurance_2016)
general_cohort_wide$FPL_Level_2016 <- factor(general_cohort_wide$FPL_Level_2016)
general_cohort_wide$HIV_Suppressed_2016 <- factor(general_cohort_wide$HIV_Suppressed_2016)
general_cohort_wide$STI_Pos_Anogenital_2016 <- factor(general_cohort_wide$STI_Pos_Anogenital_2016)

general_cohort_wide$Insurance_2017 <- factor(general_cohort_wide$Insurance_2017)
general_cohort_wide$FPL_Level_2017 <- factor(general_cohort_wide$FPL_Level_2017)
general_cohort_wide$HIV_Suppressed_2017 <- factor(general_cohort_wide$HIV_Suppressed_2017)
general_cohort_wide$STI_Pos_Anogenital_2017 <- factor(general_cohort_wide$STI_Pos_Anogenital_2017)

general_cohort_wide$Insurance_2018 <- factor(general_cohort_wide$Insurance_2018)
general_cohort_wide$FPL_Level_2018 <- factor(general_cohort_wide$FPL_Level_2018)
general_cohort_wide$HIV_Suppressed_2018 <- factor(general_cohort_wide$HIV_Suppressed_2018)
general_cohort_wide$STI_Pos_Anogenital_2018 <- factor(general_cohort_wide$STI_Pos_Anogenital_2018)

general_cohort_wide$Insurance_2019 <- factor(general_cohort_wide$Insurance_2019)
general_cohort_wide$FPL_Level_2019 <- factor(general_cohort_wide$FPL_Level_2019)
general_cohort_wide$HIV_Suppressed_2019 <- factor(general_cohort_wide$HIV_Suppressed_2019)
general_cohort_wide$STI_Pos_Anogenital_2019 <- factor(general_cohort_wide$STI_Pos_Anogenital_2019)
```

## Multiple Imputation with Chained Equations

```{r}
#Establishing imputation methods and an initial predictor matrix for the final imputation model

library(mice)

init = mice(general_cohort_wide, maxit=0, seed=0829) 
meth = init$method

meth[c("ID")]=""
meth[c("Site")]=""
meth[c("Age")]="pmm"
meth[c("Gender")]="polyreg"
meth[c("Race")]="polyreg"
meth[c("Ethnicity")]="cart"

meth[c("Insurance_2016")]="polyreg"
meth[c("FPL_Level_2016")]="cart"
meth[c("On_Hormones_2016")]="cart"
meth[c("On_PrEP_2016")]="cart"
meth[c("HIV_Status_2016")]="cart"
meth[c("HIV_Suppressed_2016")]="cart"
meth[c("STI_Pos_Anogenital_2016")]="cart"

meth[c("Insurance_2017")]="polyreg"
meth[c("FPL_Level_2017")]="cart"
meth[c("On_Hormones_2017")]="cart"
meth[c("On_PrEP_2017")]="cart"
meth[c("HIV_Status_2017")]="cart"
meth[c("HIV_Suppressed_2017")]="cart"
meth[c("STI_Pos_Anogenital_2017")]="cart"

meth[c("Insurance_2018")]="polyreg"
meth[c("FPL_Level_2018")]="cart"
meth[c("On_Hormones_2018")]="cart"
meth[c("On_PrEP_2018")]="cart"
meth[c("HIV_Status_2018")]="cart"
meth[c("HIV_Suppressed_2018")]="cart"
meth[c("STI_Pos_Anogenital_2018")]="cart"

meth[c("Insurance_2019")]="polyreg"
meth[c("FPL_Level_2019")]="cart"
meth[c("On_Hormones_2019")]="cart"
meth[c("On_PrEP_2019")]="cart"
meth[c("HIV_Status_2019")]="cart"
meth[c("HIV_Suppressed_2019")]="cart"
meth[c("STI_Pos_Anogenital_2019")]="cart"

predM = init$predictorMatrix
```

```{r}
#Editing the predictor matrix to remove IDs and post-baseline outcomes from the imputation model
predM[,1] <- 0
predM[,14:34] <- 0
```

```{r, eval = FALSE, echo = FALSE}
#Applying multiple imputation with chained equations to the wide cohort dataset.
library(mice)

imputed_data <- mice(general_cohort_wide, method=meth, predictorMatrix = predM, m=100, seed = 0829)
```

```{r}
wide_imputed_dataset <- complete(imputed_data, action="long", include=FALSE)
```

```{r}
library(miceadds)

#Checking missingness in the imputed datasets
data.frame(prop_miss(wide_imputed_dataset))
```

```{r}
#Examining raw cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$On_PrEP_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$On_PrEP_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$On_PrEP_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$On_PrEP_2019)
```

```{r}
#Recoding on PrEP variable to "No" for folks who are HIV+
wide_imputed_dataset$On_PrEP_2016[wide_imputed_dataset$HIV_Status_2016 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2017[wide_imputed_dataset$HIV_Status_2017 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2018[wide_imputed_dataset$HIV_Status_2018 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2019[wide_imputed_dataset$HIV_Status_2019 == "HIV+"] <- "No"
```

```{r}
#Examining cleaned cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$On_PrEP_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$On_PrEP_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$On_PrEP_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$On_PrEP_2019)
```

```{r}
#Examining raw cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$HIV_Suppressed_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$HIV_Suppressed_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$HIV_Suppressed_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$HIV_Suppressed_2019)
```

```{r}
#Recoding viral suppression variable to "<200 copies" for folks who are HIV-
wide_imputed_dataset$HIV_Suppressed_2016[wide_imputed_dataset$HIV_Status_2016 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2017[wide_imputed_dataset$HIV_Status_2017 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2018[wide_imputed_dataset$HIV_Status_2018 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2019[wide_imputed_dataset$HIV_Status_2019 == "HIV-"] <- "<200 copies"
```

```{r}
#Examining cleaned cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$HIV_Suppressed_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$HIV_Suppressed_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$HIV_Suppressed_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$HIV_Suppressed_2019)
```

## Sociodemographics for the Imputed Dataset

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2016 = case_when(
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "No") ~ 0,
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "Yes") ~ 1,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "<200 copies") ~ 2,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2016 <- factor(wide_imputed_dataset$HIV_Risk_2016)

levels(wide_imputed_dataset$HIV_Risk_2016) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2017 = case_when(
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "No") ~ 0,
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "Yes") ~ 1,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "<200 copies") ~ 2,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2017 <- factor(wide_imputed_dataset$HIV_Risk_2017)

levels(wide_imputed_dataset$HIV_Risk_2017) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2018 = case_when(
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "No") ~ 0,
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "Yes") ~ 1,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "<200 copies") ~ 2,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2018 <- factor(wide_imputed_dataset$HIV_Risk_2018)

levels(wide_imputed_dataset$HIV_Risk_2018) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2019 = case_when(
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "No") ~ 0,
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "Yes") ~ 1,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "<200 copies") ~ 2,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2019 <- factor(wide_imputed_dataset$HIV_Risk_2019)

levels(wide_imputed_dataset$HIV_Risk_2019) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
table(wide_imputed_dataset$HIV_Risk_2016)
table(wide_imputed_dataset$HIV_Risk_2017)
table(wide_imputed_dataset$HIV_Risk_2018)
table(wide_imputed_dataset$HIV_Risk_2019)
```

```{r}
#Creating a categorical variable for age
wide_imputed_dataset <- mutate(wide_imputed_dataset, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

wide_imputed_dataset$AgeCat <- factor(wide_imputed_dataset$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))

```

```{r}
library(stats)

median(wide_imputed_dataset$Age)
IQR(wide_imputed_dataset$Age)
```

```{r}
#Generating a descriptive statistic table for the imputed dataset at baseline (2016)

library(tableone)

imputed_baseline_table <- CreateTableOne(vars = c("Site", "Age", "AgeCat", "Gender", "Race", "Ethnicity", 
                                                  "Insurance_2016", "FPL_Level_2016", "On_Hormones_2016", 
                                                  "On_PrEP_2016", "HIV_Status_2016", "HIV_Suppressed_2016", 
                                                  "HIV_Risk_2016", "STI_Pos_Anogenital_2016"),
                                         data = wide_imputed_dataset, includeNA = TRUE)

imputed_baseline_table
```

```{r}
#Comparing to observed data at baseline

general_baseline_table
```

## Creating Imputed Repeated Measures Dataset

```{r}
#Creating placeholder variables for imputation and ID keys
wide_imputed_dataset$Imp_Num <- wide_imputed_dataset$".imp"
wide_imputed_dataset$Imp_ID <- wide_imputed_dataset$".id"
```

```{r}
impute_data_2016 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, Insurance_2016 as Insurance, FPL_Level_2016 as FPL_Level, On_Hormones_2016 as On_Hormones, On_PrEP_2016 as On_PrEP, HIV_Status_2016 as HIV_Status,
HIV_Suppressed_2016 as HIV_Suppressed, STI_Pos_Anogenital_2016 as STI_Pos_Anogenital FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2016$Year <- 2016
```

```{r}
impute_data_2017 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, Insurance_2017 as Insurance, FPL_Level_2017 as FPL_Level, On_Hormones_2017 as On_Hormones, On_PrEP_2017 as On_PrEP, HIV_Status_2017 as HIV_Status,
HIV_Suppressed_2017 as HIV_Suppressed, STI_Pos_Anogenital_2017 as STI_Pos_Anogenital FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2017$Year <- 2017
```

```{r}
impute_data_2018 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, Insurance_2018 as Insurance, FPL_Level_2018 as FPL_Level, On_Hormones_2018 as On_Hormones, On_PrEP_2018 as On_PrEP, HIV_Status_2018 as HIV_Status,
HIV_Suppressed_2018 as HIV_Suppressed, STI_Pos_Anogenital_2018 as STI_Pos_Anogenital FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2018$Year <- 2018
```

```{r}
impute_data_2019 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, Insurance_2019 as Insurance, FPL_Level_2019 as FPL_Level, On_Hormones_2019 as On_Hormones, On_PrEP_2019 as On_PrEP, HIV_Status_2019 as HIV_Status,
HIV_Suppressed_2019 as HIV_Suppressed, STI_Pos_Anogenital_2019 as STI_Pos_Anogenital FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2019$Year <- 2019
```

```{r}
long_imputed_data <- rbind(impute_data_2016, impute_data_2017, impute_data_2018, impute_data_2019)
```

```{r}
#Recoding imputation and ID keys
long_imputed_data$".imp" <- long_imputed_data$Imp_Num
long_imputed_data$".id" <- long_imputed_data$Imp_ID

#Dropping placeholder keys
long_imputed_data <- subset(long_imputed_data, select = -c(Imp_Num, Imp_ID))
```

```{r}
long_imputed_data <- long_imputed_data[order(long_imputed_data$.imp, long_imputed_data$.id),]
```

## Cleaning Long Imputed Dataset

```{r}
#Recoding character variables as factors
long_imputed_data$Race <- factor(long_imputed_data$Race)
levels(long_imputed_data$Race)

long_imputed_data$Insurance <- factor(long_imputed_data$Insurance)
levels(long_imputed_data$Insurance)

long_imputed_data$FPL_Level <- factor(long_imputed_data$FPL_Level)
levels(long_imputed_data$FPL_Level)

long_imputed_data$On_Hormones <- factor(long_imputed_data$On_Hormones)
levels(long_imputed_data$On_Hormones)

long_imputed_data$On_PrEP <- factor(long_imputed_data$On_PrEP)
levels(long_imputed_data$On_PrEP)

long_imputed_data$HIV_Status <- factor(long_imputed_data$HIV_Status)
levels(long_imputed_data$HIV_Status)

long_imputed_data$HIV_Suppressed <- factor(long_imputed_data$HIV_Suppressed)
levels(long_imputed_data$HIV_Suppressed)

long_imputed_data$STI_Pos_Anogenital <- factor(long_imputed_data$STI_Pos_Anogenital)
levels(long_imputed_data$STI_Pos_Anogenital)
```

```{r}
#Adding a variable for observation year
long_imputed_data <- mutate(long_imputed_data, Obs_Year = case_when(
  Year == 2016 ~ 0,
  Year == 2017 ~ 1,
  Year == 2018 ~ 2,
  Year == 2019 ~ 3
))

#Creating a categorical variable for age
long_imputed_data <- mutate(long_imputed_data, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

long_imputed_data$AgeCat <- factor(long_imputed_data$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))

long_imputed_data <- mutate(long_imputed_data, RaceID = case_when(
  Race == "White" ~ 0,
  Race == "Black or African American" ~ 1,
  Race == "Multiracial" ~ 2,
  Race == "Another Race" | Race == "Asian" | Race == "American Indian or Alaska Native" | 
    Race == "Native Hawaiian or Other Pacific Islander" ~ 3
))

long_imputed_data$RaceID <- factor(long_imputed_data$RaceID)

levels(long_imputed_data$RaceID) <- c("White", "Black or African American", "Multiracial", "Another Race")


long_imputed_data <- mutate(long_imputed_data, HIV_Risk = case_when(
  (HIV_Status == "HIV-") & (On_PrEP == "No") ~ 0,
  (HIV_Status == "HIV-") & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "200+ copies") ~ 3
))

long_imputed_data$HIV_Risk <- factor(long_imputed_data$HIV_Risk)

levels(long_imputed_data$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")

```

```{r}
long_imputed_data <- mutate(long_imputed_data, STI_Positive = case_when(
  STI_Pos_Anogenital == "No" ~ 0,
  STI_Pos_Anogenital == "Yes" ~ 1
))
```

## Creating Dataset List

```{r}
imputed_data_list <- split(long_imputed_data, f = long_imputed_data$.imp)
```

## GEE Modeling Across Datasets

```{r}
library(miceadds)

imputed_mids <- datalist2mids(imputed_data_list)
```

```{r}
library(mice)
library(geepack)

#Unadjusted log-Poisson for positive anogenital STI test

STI_unadjusted <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + Site + Obs_Year, family = poisson(link = "log"), 
                                                   id = ID, waves = Obs_Year, corstr = "exchangeable"))

```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_unadjusted <- pool(STI_unadjusted, dfcom = CC_STI_unadjusted_df)

options(digits = 3)

summary(STI_combined_unadjusted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

STI_adjusted <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + Site + AgeCat + relevel(Gender, ref = "Male") + relevel(RaceID, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+")
                                                 + HIV_Risk + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_adjusted <- pool(STI_adjusted, dfcom = CC_STI_adjusted_df)

options(digits = 3)

summary(STI_combined_adjusted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

## Sensitivity Analyses

```{r}
library(EValue)

evalues.RR(est = 0.72, lo = 0.55, hi = 0.93)

bias_plot(0.75, xmax = 10)
```

## EMM Analyses

```{r}
library(mice)
library(geepack)

STI_age <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + AgeCat + On_Hormones*AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_age <- pool(STI_age)

options(digits = 3)

summary(STI_combined_age, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

STI_race <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + relevel(Race, ref = "White") + On_Hormones*relevel(Race, ref = "White") + 
                                               Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_race <- pool(STI_race)

options(digits = 3)

summary(STI_combined_race, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

STI_ethnicity <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + relevel(Ethnicity, ref = "Not Hispanic") + 
                                                    On_Hormones*relevel(Ethnicity, ref = "Not Hispanic") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_ethnicity <- pool(STI_ethnicity)

options(digits = 3)

summary(STI_combined_ethnicity, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

STI_gender <- with(imputed_mids, exp <- geeglm(STI_Positive ~ On_Hormones + relevel(Gender, ref = "Male") + On_Hormones*relevel(Gender, ref = "Male") 
                                               + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
library(mice)

#Pooling estimates across datasets

STI_combined_gender <- pool(STI_gender)

options(digits = 3)

summary(STI_combined_gender, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```


