---
title: "PHQ EMR - LEGACY DFRR"
author: "David R. Pletta, MPH"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Formatting output for knitr
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE)

#Setting the working directory for the markdown document
knitr::opts_knit$set(root.dir = "/Users/davidpletta/Library/CloudStorage/Dropbox/ Data Projects (David Pletta)/LEGACY/Reporting - DFRR/DFRR PCORI/AIM II - GA Care and Mental Health/PHQ EMR")
```

```{r}
library(tidyverse)
library(tidylog)
```

```{r}
#Importing EMR cohorts from R dataset
load("EMR Cohorts_LEGACY_07202022.RData")
```

### Cleaning Sociodemographics

```{r}
#Removing age observations above 100 years old
general_cohort$Age[general_cohort$Age > 100] <- NA
```

```{r}
table(general_cohort$Age, useNA = "always")
```

```{r}
#Creating a categorical variable for age
general_cohort <- mutate(general_cohort, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

general_cohort$AgeCat <- factor(general_cohort$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))

```

```{r}
general_cohort$On_PrEP <- factor(general_cohort$On_PrEP)

levels(general_cohort$On_PrEP) <- c("No", "Yes")

table(general_cohort$On_PrEP, general_cohort$Year)
```

```{r}
general_cohort <- mutate(general_cohort, Gender = case_when(
  GenderID == "Female" ~ 0,
  GenderID == "Male" ~ 1,
  (GenderID == "Nonbinary") & (Sex_at_Birth == "Female") ~ 2,
  (GenderID == "Nonbinary") & (Sex_at_Birth == "Male") ~ 3
))

general_cohort$Gender <- factor(general_cohort$Gender)

levels(general_cohort$Gender) <- c("Female", "Male", "Nonbinary AFAB", "Nonbinary AMAB")

table(general_cohort$Gender)

prop.table(table(general_cohort$Gender))
```

```{r}
general_cohort <- mutate(general_cohort, HIV_Risk = case_when(
  (HIV_Status == 0) & (On_PrEP == "No") ~ 0,
  (HIV_Status == 0) & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == 1) & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == 1) & (HIV_Suppressed == "200+ copies") ~ 3
))

general_cohort$HIV_Risk <- factor(general_cohort$HIV_Risk)

levels(general_cohort$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
table(general_cohort$HIV_Risk)

prop.table(table(general_cohort$HIV_Risk))
```

```{r}
general_cohort$HIV_Status <- factor(general_cohort$HIV_Status)

levels(general_cohort$HIV_Status) <- c("HIV-", "HIV+")

table(general_cohort$HIV_Status)
```

```{r}
general_cohort$PHQ_Clinical <- factor(general_cohort$PHQ_Clinical)

levels(general_cohort$PHQ_Clinical) <- c("No", "Yes")

table(general_cohort$PHQ_Clinical, useNA = "always")
```

## Calculating Years in Cohort

```{r}
library(sqldf)

cohort_years_data <- sqldf('SELECT ID, count(ID) as cohort_years FROM general_cohort GROUP BY ID ORDER BY ID')
```

```{r}
table(cohort_years_data$cohort_years)
```

```{r}
general_cohort_revised <- sqldf('SELECT a.*, b.cohort_years FROM general_cohort as a LEFT JOIN cohort_years_data as b ON a.ID = b.ID')
```

```{r}
general_cohort <- general_cohort_revised
```

## Restricting to Those with PHQ Data

```{r}
table(general_cohort$PHQ_Clinical, useNA = "always")
```

```{r}
general_cohort_restricted <- general_cohort[which(!is.na(general_cohort$PHQ_Clinical)),]
```

```{r}
table(general_cohort_restricted$Site)
```

## Examining Missingness in Restricted Cohort

```{r}
library(tidyverse)
library(dplyr)

general_cohort_restricted_impmodel <- general_cohort_restricted[c("ID", "Year", "Site", "AgeCat", "Gender", "Race", "Ethnicity", "Insurance", "FPL_Level",
                                                                 "On_Hormones", "On_PrEP", "HIV_Status", "HIV_Suppressed", "cohort_years", "PHQ_Clinical")]

general_cohort_restricted_impmodel_2016 <- general_cohort_restricted_impmodel[which(general_cohort_restricted_impmodel$Year == 2016),]
```

```{r}
library(miceadds)

#Checking missingness in the imputed datasets
data.frame(prop_miss(general_cohort_restricted_impmodel_2016))
```

```{r}
library(tidyverse)
library(dplyr)

general_cohort_restricted_impmodel_2 <- general_cohort_restricted[c("ID", "Year", "Site", "AgeCat", "Gender", "Race", "Ethnicity", "Insurance", "FPL_Level",
                                                                 "On_Hormones", "HIV_Status", "cohort_years", "PHQ_Clinical")]

general_cohort_restricted_impmodel_2016_2 <- general_cohort_restricted_impmodel_2[which(general_cohort_restricted_impmodel_2$Year == 2016),]
```

```{r}
library(miceadds)

#Checking missingness in the imputed datasets
data.frame(prop_miss(general_cohort_restricted_impmodel_2016_2))
```

```{r}
general_cohort_restricted_impmodel_2016_2$n_miss <- rowSums(is.na(general_cohort_restricted_impmodel_2016_2))

general_cohort_restricted_impmodel_2016_2$any_miss <- ifelse(general_cohort_restricted_impmodel_2016_2$n_miss >= 1, 1, 0)

table(general_cohort_restricted_impmodel_2016_2$any_miss)
```





## Attrition Analysis

```{r}
library(sqldf)

baseline_PHQ <- sqldf('SELECT ID, PHQ_Clinical as PHQ_Clinical_2016 FROM general_cohort_restricted WHERE Year = 2016 ORDER BY ID')
```

```{r}
library(sqldf)

obsyear_data <- sqldf('SELECT ID, count(ID) as obs_years FROM general_cohort_restricted GROUP BY ID ORDER BY ID')
```

```{r}
library(sqldf)

attrition_data <- sqldf('SELECT a.ID, a.PHQ_Clinical_2016, b.obs_years FROM baseline_PHQ as a LEFT JOIN obsyear_data as b ON a.ID = b.ID ORDER BY a.ID')
```

```{r}
t.test(attrition_data$obs_years ~ attrition_data$PHQ_Clinical_2016)
```


```{r}
attrition_model <- glm(obs_years ~ PHQ_Clinical_2016, data = attrition_data)
summary(attrition_model)
```


## Cohort Sociodemographics at Baseline (2016)

```{r}
#Subsetting baseline data for the general cohort
general_baseline_restricted <- general_cohort_restricted[which(general_cohort_restricted$Year == 2016),]
```

```{r}
library(stats)

median(general_baseline_restricted$Age, na.rm = TRUE)
IQR(general_baseline_restricted$Age, na.rm = TRUE)
```

```{r}
library(tableone)

general_baseline_restricted_table <- CreateTableOne(vars = c("Site", "AgeCat","Gender", "Race", "Ethnicity", 
                                                  "Insurance", "FPL_Level", "On_Hormones", 
                                                  "HIV_Risk", "cohort_years"),
                                         strata = "PHQ_Clinical",
                                         data = general_baseline_restricted, includeNA = TRUE,
                                         addOverall = TRUE)

general_baseline_restricted_table
```

### Complete Case Model

```{r}
general_cohort_2016 <- general_cohort[which(general_cohort$Year == 2016),]

table(general_cohort_2016$PHQ_Clinical, useNA = 'always')

baseline_phq <- general_cohort_2016$ID[!is.na(general_cohort_2016$PHQ_Clinical)]
```

```{r}
general_cohort_complete_temp <- subset(general_cohort_restricted, select = c(ID, Year, Site, AgeCat, Gender, Race, Ethnicity, Insurance, FPL_Level,
                                                                 On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, cohort_years, PHQ_Clinical))

#general_cohort_completecase <- general_cohort_complete_temp[which(general_cohort_complete_temp$ID %in% baseline_phq),]

#Creating a complete-case dataset
general_cohort_completecase <- general_cohort_complete_temp[complete.cases(general_cohort_complete_temp),]
```

```{r}
general_cohort_completecase <- mutate(general_cohort_completecase, HIV_Risk = case_when(
  (HIV_Status == "HIV-") & (On_PrEP == "No") ~ 0,
  (HIV_Status == "HIV-") & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "200+ copies") ~ 3
))

general_cohort_completecase$HIV_Risk <- factor(general_cohort_completecase$HIV_Risk)

levels(general_cohort_completecase$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")

general_cohort_completecase <- mutate(general_cohort_completecase, Obs_Year = case_when(
  Year == 2016 ~ 0,
  Year == 2017 ~ 1,
  Year == 2018 ~ 2,
  Year == 2019 ~ 3
))
```

```{r}
#Collapsing race categories

general_cohort_completecase <- mutate(general_cohort_completecase, Race_Three = case_when(
  Race == "White" ~ 0,
  Race == "Asian" ~ 1,
  Race == "Black or African American" ~ 2,
  Race == "American Indian or Alaska Native" | Race == "Another Race" | Race == "Multiracial" | Race == "Native Hawaiian or Other Pacific Islander" ~ 3
))

general_cohort_completecase$Race_Three <- factor(general_cohort_completecase$Race_Three)
levels(general_cohort_completecase$Race_Three) <- c("White", "Asian", "Black/African American", "Another race")

table(general_cohort_completecase$Race_Three)
```

```{r}
general_cohort_completecase$PHQ_numeric <- NA
general_cohort_completecase$PHQ_numeric[general_cohort_completecase$PHQ_Clinical == "No"] <- 0
general_cohort_completecase$PHQ_numeric[general_cohort_completecase$PHQ_Clinical == "Yes"] <- 1
```

```{r}
general_cohort_completecase$Gender <- factor(general_cohort_completecase$Gender)
general_cohort_completecase$Ethnicity <- factor(general_cohort_completecase$Ethnicity)
general_cohort_completecase$Insurance <- factor(general_cohort_completecase$Insurance)
general_cohort_completecase$FPL_Level <- factor(general_cohort_completecase$FPL_Level)
```

```{r}
table(general_cohort_completecase$Gender)
```

```{r}
general_cohort_completecase <- mutate(general_cohort_completecase, Gender_Three = case_when(
  Gender == "Male" ~ 0,
  Gender == "Female" ~ 1,
  Gender == "Nonbinary AFAB" | Gender == "Nonbinary AMAB" ~ 2
))

general_cohort_completecase$Gender_Three <- factor(general_cohort_completecase$Gender_Three)
levels(general_cohort_completecase$Gender_Three) <- c("Male", "Female", "Nonbinary")
table(general_cohort_completecase$Gender_Three)
```


```{r}
library(geepack)

CC_PHQ_adjusted <- geeglm(PHQ_numeric ~ On_Hormones + AgeCat + relevel(Gender_Three, ref = "Male") + relevel(Race_Three, ref = "White") + relevel(Ethnicity, ref = "Not Hispanic") + relevel(Insurance, ref = "Private insurance") + relevel(FPL_Level, ref = "251%+") + HIV_Risk + cohort_years + Obs_Year, 
                          family = poisson(link = "log"), data = general_cohort_completecase, id = ID, waves = Obs_Year, corstr = "exchangeable")

options(scipen = 5)

tidy(CC_PHQ_adjusted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
CC_PHQ_adjusted_df <- df.residual(CC_PHQ_adjusted)
```

### Creating Wide Dataset for Imputation

```{r}
#Creating separate datasets for baseline sociodemographics & time-varying demographics and outcomes
library(sqldf)

Cohort_Demographics <- sqldf('SELECT ID, Site, Age, Gender, 
                             Race, Ethnicity, cohort_years FROM general_cohort WHERE Year = 2016 ORDER BY ID')

Outcomes_2016 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical 
                       FROM general_cohort WHERE Year = 2016 ORDER BY ID')

Outcomes_2017 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical 
                       FROM general_cohort WHERE Year = 2017 ORDER BY ID')

Outcomes_2018 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical
                       FROM general_cohort WHERE Year = 2018 ORDER BY ID')

Outcomes_2019 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical
                       FROM general_cohort WHERE Year = 2019 ORDER BY ID')
```

```{r, message = FALSE, echo = FALSE}
#Renaming variables in the outcome datasets before proceeding to left joins.
library(data.table)

setnames(Outcomes_2016, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2016', 'FPL_Level_2016', 'On_Hormones_2016', 'On_PrEP_2016', 
                                'HIV_Status_2016', 'HIV_Suppressed_2016', 'PHQ_Clinical_2016'))

setnames(Outcomes_2017, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2017', 'FPL_Level_2017', 'On_Hormones_2017', 'On_PrEP_2017', 
                                'HIV_Status_2017', 'HIV_Suppressed_2017', 'PHQ_Clinical_2017'))

setnames(Outcomes_2018, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2018', 'FPL_Level_2018', 'On_Hormones_2018', 'On_PrEP_2018', 
                                'HIV_Status_2018', 'HIV_Suppressed_2018', 'PHQ_Clinical_2018'))

setnames(Outcomes_2019, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2019', 'FPL_Level_2019', 'On_Hormones_2019', 'On_PrEP_2019', 
                                'HIV_Status_2019', 'HIV_Suppressed_2019', 'PHQ_Clinical_2019'))

```

```{r}
#Left joining the datasets on patient ID

Demo_16 <- merge(Cohort_Demographics, Outcomes_2016, by = "ID")

Demo_1617 <- merge(Demo_16, Outcomes_2017, by = "ID", all.x = TRUE)

Demo_161718 <- merge(Demo_1617, Outcomes_2018, by = "ID", all.x = TRUE)

general_cohort_wide <- merge(Demo_161718, Outcomes_2019, by = "ID", all.x = TRUE)
```

```{r}
#Converting character variables to factor variables
general_cohort_wide$Gender <- factor(general_cohort_wide$Gender)
general_cohort_wide$Race <- factor(general_cohort_wide$Race)
general_cohort_wide$Ethnicity <- factor(general_cohort_wide$Ethnicity)

general_cohort_wide$Insurance_2016 <- factor(general_cohort_wide$Insurance_2016)
general_cohort_wide$FPL_Level_2016 <- factor(general_cohort_wide$FPL_Level_2016)
general_cohort_wide$HIV_Suppressed_2016 <- factor(general_cohort_wide$HIV_Suppressed_2016)
general_cohort_wide$PHQ_Clinical_2016 <- factor(general_cohort_wide$PHQ_Clinical_2016)

general_cohort_wide$Insurance_2017 <- factor(general_cohort_wide$Insurance_2017)
general_cohort_wide$FPL_Level_2017 <- factor(general_cohort_wide$FPL_Level_2017)
general_cohort_wide$HIV_Suppressed_2017 <- factor(general_cohort_wide$HIV_Suppressed_2017)
general_cohort_wide$PHQ_Clinical_2017 <- factor(general_cohort_wide$PHQ_Clinical_2017)

general_cohort_wide$Insurance_2018 <- factor(general_cohort_wide$Insurance_2018)
general_cohort_wide$FPL_Level_2018 <- factor(general_cohort_wide$FPL_Level_2018)
general_cohort_wide$HIV_Suppressed_2018 <- factor(general_cohort_wide$HIV_Suppressed_2018)
general_cohort_wide$PHQ_Clinical_2018 <- factor(general_cohort_wide$PHQ_Clinical_2018)

general_cohort_wide$Insurance_2019 <- factor(general_cohort_wide$Insurance_2019)
general_cohort_wide$FPL_Level_2019 <- factor(general_cohort_wide$FPL_Level_2019)
general_cohort_wide$HIV_Suppressed_2019 <- factor(general_cohort_wide$HIV_Suppressed_2019)
general_cohort_wide$PHQ_Clinical_2019 <- factor(general_cohort_wide$PHQ_Clinical_2019)
```

### Multiple Imputation with Chained Equations

```{r}
#Establishing imputation methods and an initial predictor matrix for the final imputation model

library(mice)

init = mice(general_cohort_wide, maxit=0, seed=0829) 
meth = init$method

meth[c("ID")]=""
meth[c("Site")]=""
meth[c("Age")]="pmm"
meth[c("Gender")]="polyreg"
meth[c("Race")]="polyreg"
meth[c("Ethnicity")]="cart"
meth[c("cohort_years")]=""

meth[c("Insurance_2016")]="polyreg"
meth[c("FPL_Level_2016")]="cart"
meth[c("On_Hormones_2016")]="cart"
meth[c("On_PrEP_2016")]="cart"
meth[c("HIV_Status_2016")]="cart"
meth[c("HIV_Suppressed_2016")]="cart"
meth[c("PHQ_Clinical_2016")]="cart"

meth[c("Insurance_2017")]="polyreg"
meth[c("FPL_Level_2017")]="cart"
meth[c("On_Hormones_2017")]="cart"
meth[c("On_PrEP_2017")]="cart"
meth[c("HIV_Status_2017")]="cart"
meth[c("HIV_Suppressed_2017")]="cart"
meth[c("PHQ_Clinical_2017")]="cart"

meth[c("Insurance_2018")]="polyreg"
meth[c("FPL_Level_2018")]="cart"
meth[c("On_Hormones_2018")]="cart"
meth[c("On_PrEP_2018")]="cart"
meth[c("HIV_Status_2018")]="cart"
meth[c("HIV_Suppressed_2018")]="cart"
meth[c("PHQ_Clinical_2018")]="cart"

meth[c("Insurance_2019")]="polyreg"
meth[c("FPL_Level_2019")]="cart"
meth[c("On_Hormones_2019")]="cart"
meth[c("On_PrEP_2019")]="cart"
meth[c("HIV_Status_2019")]="cart"
meth[c("HIV_Suppressed_2019")]="cart"
meth[c("PHQ_Clinical_2019")]="cart"

predM = init$predictorMatrix
```

```{r}
#Editing the predictor matrix to remove IDs and post-baseline outcomes from the imputation model
predM[,1] <- 0
predM[,15:35] <- 0
```

```{r, echo = FALSE, eval = FALSE}
#Applying multiple imputation with chained equations to the wide cohort dataset.
library(mice)

imputed_data <- mice(general_cohort_wide, method=meth, predictorMatrix = predM, m=5, seed = 0829)
```

```{r}
wide_imputed_dataset <- complete(imputed_data, action="long", include=FALSE)
```

```{r}
library(miceadds)

#Checking missingness in the imputed datasets
data.frame(prop_miss(wide_imputed_dataset))
```

### Cleaning Imputed Dataset

```{r}
#Examining raw cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$On_PrEP_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$On_PrEP_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$On_PrEP_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$On_PrEP_2019)
```

```{r}
#Recoding on PrEP variable to "No" for folks who are HIV+
wide_imputed_dataset$On_PrEP_2016[wide_imputed_dataset$HIV_Status_2016 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2017[wide_imputed_dataset$HIV_Status_2017 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2018[wide_imputed_dataset$HIV_Status_2018 == "HIV+"] <- "No"
wide_imputed_dataset$On_PrEP_2019[wide_imputed_dataset$HIV_Status_2019 == "HIV+"] <- "No"
```

```{r}
#Examining cleaned cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$On_PrEP_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$On_PrEP_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$On_PrEP_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$On_PrEP_2019)
```

```{r}
#Examining raw cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$HIV_Suppressed_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$HIV_Suppressed_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$HIV_Suppressed_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$HIV_Suppressed_2019)
```

```{r}
#Recoding viral suppression variable to "<200 copies" for folks who are HIV-
wide_imputed_dataset$HIV_Suppressed_2016[wide_imputed_dataset$HIV_Status_2016 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2017[wide_imputed_dataset$HIV_Status_2017 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2018[wide_imputed_dataset$HIV_Status_2018 == "HIV-"] <- "<200 copies"
wide_imputed_dataset$HIV_Suppressed_2019[wide_imputed_dataset$HIV_Status_2019 == "HIV-"] <- "<200 copies"
```

```{r}
#Examining cleaned cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset$HIV_Status_2016, wide_imputed_dataset$HIV_Suppressed_2016)
table(wide_imputed_dataset$HIV_Status_2017, wide_imputed_dataset$HIV_Suppressed_2017)
table(wide_imputed_dataset$HIV_Status_2018, wide_imputed_dataset$HIV_Suppressed_2018)
table(wide_imputed_dataset$HIV_Status_2019, wide_imputed_dataset$HIV_Suppressed_2019)
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2016 = case_when(
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "No") ~ 0,
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "Yes") ~ 1,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "<200 copies") ~ 2,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2016 <- factor(wide_imputed_dataset$HIV_Risk_2016)

levels(wide_imputed_dataset$HIV_Risk_2016) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2017 = case_when(
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "No") ~ 0,
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "Yes") ~ 1,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "<200 copies") ~ 2,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2017 <- factor(wide_imputed_dataset$HIV_Risk_2017)

levels(wide_imputed_dataset$HIV_Risk_2017) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2018 = case_when(
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "No") ~ 0,
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "Yes") ~ 1,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "<200 copies") ~ 2,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2018 <- factor(wide_imputed_dataset$HIV_Risk_2018)

levels(wide_imputed_dataset$HIV_Risk_2018) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset <- mutate(wide_imputed_dataset, HIV_Risk_2019 = case_when(
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "No") ~ 0,
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "Yes") ~ 1,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "<200 copies") ~ 2,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "200+ copies") ~ 3
))

wide_imputed_dataset$HIV_Risk_2019 <- factor(wide_imputed_dataset$HIV_Risk_2019)

levels(wide_imputed_dataset$HIV_Risk_2019) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
table(wide_imputed_dataset$HIV_Risk_2016)
table(wide_imputed_dataset$HIV_Risk_2017)
table(wide_imputed_dataset$HIV_Risk_2018)
table(wide_imputed_dataset$HIV_Risk_2019)
```

### Sociodemographics at Baseline for Imputed Dataset

```{r}
#Creating a categorical variable for age
wide_imputed_dataset <- mutate(wide_imputed_dataset, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

wide_imputed_dataset$AgeCat <- factor(wide_imputed_dataset$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))
```

```{r}
#Generating a descriptive statistic table for the general imputed dataset at baseline (2016) for the PHQ analysis

library(tableone)

imputed_baseline_table <- CreateTableOne(vars = c("Site", "Age", "AgeCat", "Gender", "Race", "Ethnicity", 
                                                  "Insurance_2016", "FPL_Level_2016", "On_Hormones_2016", 
                                                  "HIV_Risk_2016"),
                                         strata = "PHQ_Clinical_2016",
                                         data = wide_imputed_dataset, includeNA = TRUE,
                                         addOverall = TRUE)

imputed_baseline_table
```


## Creating Imputed Repeated Measures Dataset

```{r}
#Creating placeholder variables for imputation and ID keys
wide_imputed_dataset$Imp_Num <- wide_imputed_dataset$".imp"
wide_imputed_dataset$Imp_ID <- wide_imputed_dataset$".id"
```

```{r}
impute_data_2016 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2016 as Insurance, FPL_Level_2016 as FPL_Level, On_Hormones_2016 as On_Hormones, On_PrEP_2016 as On_PrEP, HIV_Status_2016 as HIV_Status,
HIV_Suppressed_2016 as HIV_Suppressed, PHQ_Clinical_2016 as PHQ_Clinical FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2016$Year <- 2016
```

```{r}
impute_data_2017 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2017 as Insurance, FPL_Level_2017 as FPL_Level, On_Hormones_2017 as On_Hormones, On_PrEP_2017 as On_PrEP, HIV_Status_2017 as HIV_Status,
HIV_Suppressed_2017 as HIV_Suppressed, PHQ_Clinical_2017 as PHQ_Clinical FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2017$Year <- 2017
```

```{r}
#Creating placeholder variables for imputation and ID keys
wide_imputed_dataset$Imp_Num <- wide_imputed_dataset$".imp"
wide_imputed_dataset$Imp_ID <- wide_imputed_dataset$".id"
```

```{r}
impute_data_2018 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2018 as Insurance, FPL_Level_2018 as FPL_Level, On_Hormones_2018 as On_Hormones, On_PrEP_2018 as On_PrEP, HIV_Status_2018 as HIV_Status,
HIV_Suppressed_2018 as HIV_Suppressed, PHQ_Clinical_2018 as PHQ_Clinical FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2018$Year <- 2018
```

```{r}
impute_data_2019 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2019 as Insurance, FPL_Level_2019 as FPL_Level, On_Hormones_2019 as On_Hormones, On_PrEP_2019 as On_PrEP, HIV_Status_2019 as HIV_Status,
HIV_Suppressed_2019 as HIV_Suppressed, PHQ_Clinical_2019 as PHQ_Clinical FROM wide_imputed_dataset ORDER BY Imp_Num")

impute_data_2019$Year <- 2019
```

```{r}
long_imputed_data <- rbind(impute_data_2016, impute_data_2017, impute_data_2018, impute_data_2019)
```

```{r}
#Adding a variable for observation year
long_imputed_data <- mutate(long_imputed_data, Obs_Year = case_when(
  Year == 2016 ~ 0,
  Year == 2017 ~ 1,
  Year == 2018 ~ 2,
  Year == 2019 ~ 3
))

#Creating a categorical variable for age
long_imputed_data <- mutate(long_imputed_data, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

long_imputed_data$AgeCat <- factor(long_imputed_data$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))
```

```{r}
#Recoding imputation and ID keys
long_imputed_data$".imp" <- long_imputed_data$Imp_Num
long_imputed_data$".id" <- long_imputed_data$Imp_ID

#Dropping placeholder keys
long_imputed_data <- subset(long_imputed_data, select = -c(Imp_Num, Imp_ID))
```

```{r}
long_imputed_data <- long_imputed_data[order(long_imputed_data$.imp, long_imputed_data$.id),]
```

## Cleaning Long Imputed Dataset

```{r}
#Recoding character variables as factors
long_imputed_data$Insurance <- factor(long_imputed_data$Insurance)
levels(long_imputed_data$Insurance)

long_imputed_data$FPL_Level <- factor(long_imputed_data$FPL_Level)
levels(long_imputed_data$FPL_Level)

long_imputed_data$On_Hormones <- factor(long_imputed_data$On_Hormones)
levels(long_imputed_data$On_Hormones)

long_imputed_data$On_PrEP <- factor(long_imputed_data$On_PrEP)
levels(long_imputed_data$On_PrEP)

long_imputed_data$HIV_Status <- factor(long_imputed_data$HIV_Status)
levels(long_imputed_data$HIV_Status)

long_imputed_data$HIV_Suppressed <- factor(long_imputed_data$HIV_Suppressed)
levels(long_imputed_data$HIV_Suppressed)

long_imputed_data$PHQ_Clinical <- factor(long_imputed_data$PHQ_Clinical)
levels(long_imputed_data$PHQ_Clinical)
```

```{r}
long_imputed_data <- mutate(long_imputed_data, HIV_Risk = case_when(
  (HIV_Status == "HIV-") & (On_PrEP == "No") ~ 0,
  (HIV_Status == "HIV-") & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "200+ copies") ~ 3
))

long_imputed_data$HIV_Risk <- factor(long_imputed_data$HIV_Risk)

levels(long_imputed_data$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")

```

```{r}
long_imputed_data <- mutate(long_imputed_data, PHQ_Numeric = case_when(
  PHQ_Clinical == "No" ~ 0,
  PHQ_Clinical == "Yes" ~ 1
))
```

## Creating Dataset List

```{r}
imputed_data_list <- split(long_imputed_data, f = long_imputed_data$.imp)
```

## GEE Modeling Across Datasets

```{r}
library(miceadds)

imputed_mids <- datalist2mids(imputed_data_list)
```


## Bivariate Models (Non-Restricted)

```{r}
library(mice)
library(geepack)

PHQ_bivariate_hormones <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hormones <- pool(PHQ_bivariate_hormones)

options(digits = 3)

summary(PHQ_combined_hormones, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_site <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ Site + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_site <- pool(PHQ_bivariate_site)

options(digits = 3)

summary(PHQ_combined_site, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_age <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_age <- pool(PHQ_bivariate_age)

options(digits = 3)

summary(PHQ_combined_age, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_gender <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ relevel(Gender, ref = "Male") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_gender <- pool(PHQ_bivariate_gender)

options(digits = 3)

summary(PHQ_combined_gender, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_race <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ relevel(Race, ref = "White") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_race <- pool(PHQ_bivariate_race)

options(digits = 3)

summary(PHQ_combined_race, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_ethnicity <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ relevel(Ethnicity, ref = "Not Hispanic") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_ethnicity <- pool(PHQ_bivariate_ethnicity)

options(digits = 3)

summary(PHQ_combined_ethnicity, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_insurance <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ Insurance + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_insurance <- pool(PHQ_bivariate_insurance)

options(digits = 3)

summary(PHQ_combined_insurance, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_fpl <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ relevel(FPL_Level, ref = "251%+") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_fpl <- pool(PHQ_bivariate_fpl)

options(digits = 3)

summary(PHQ_combined_fpl, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_hiv <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ HIV_Risk + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hiv <- pool(PHQ_bivariate_hiv)

options(digits = 3)

summary(PHQ_combined_hiv, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```
```{r}
library(mice)
library(geepack)

PHQ_bivariate_cohortyears <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_cohortyears <- pool(PHQ_bivariate_cohortyears)

options(digits = 3)

summary(PHQ_combined_cohortyears, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

## Multivariable Model (Non-Restricted)

```{r}
library(mice)
library(geepack)

PHQ_adjusted <- with(imputed_mids, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Site + AgeCat + relevel(Gender, ref = "Male") + relevel(Race, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+")
                                                 + HIV_Risk + cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_adjusted <- pool(PHQ_adjusted)

options(digits = 3)

summary(PHQ_combined_adjusted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```


## E-value

```{r}
library(EValue)

evalues.RR(est = 0.854, lo = 0.746, hi = 0.977)
```



# Restricted Sample Imputation and Analyses

```{r}
table(general_cohort_restricted$Year)
```


```{r}
table(general_cohort_restricted$Year, general_cohort_restricted$On_Hormones)
```
```{r}
table(general_cohort_restricted$Year, general_cohort_restricted$PHQ_Clinical)
```



### Creating Wide Dataset for Imputation

```{r}
#Creating separate datasets for baseline sociodemographics & time-varying demographics and outcomes
library(sqldf)

Cohort_Demographics_Restricted <- sqldf('SELECT ID, Site, Age, Gender, 
                             Race, Ethnicity, cohort_years FROM general_cohort_restricted WHERE Year = 2016 ORDER BY ID')

Outcomes_Restricted_2016 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical 
                       FROM general_cohort_restricted WHERE Year = 2016 ORDER BY ID')

Outcomes_Restricted_2017 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical 
                       FROM general_cohort_restricted WHERE Year = 2017 ORDER BY ID')

Outcomes_Restricted_2018 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical
                       FROM general_cohort_restricted WHERE Year = 2018 ORDER BY ID')

Outcomes_Restricted_2019 <- sqldf('SELECT ID, Insurance, FPL_Level, On_Hormones, On_PrEP, HIV_Status, HIV_Suppressed, PHQ_Clinical
                       FROM general_cohort_restricted WHERE Year = 2019 ORDER BY ID')
```

```{r, message = FALSE, echo = FALSE}
#Renaming variables in the outcome datasets before proceeding to left joins.
library(data.table)

setnames(Outcomes_Restricted_2016, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2016', 'FPL_Level_2016', 'On_Hormones_2016', 'On_PrEP_2016', 
                                'HIV_Status_2016', 'HIV_Suppressed_2016', 'PHQ_Clinical_2016'))

setnames(Outcomes_Restricted_2017, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2017', 'FPL_Level_2017', 'On_Hormones_2017', 'On_PrEP_2017', 
                                'HIV_Status_2017', 'HIV_Suppressed_2017', 'PHQ_Clinical_2017'))

setnames(Outcomes_Restricted_2018, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2018', 'FPL_Level_2018', 'On_Hormones_2018', 'On_PrEP_2018', 
                                'HIV_Status_2018', 'HIV_Suppressed_2018', 'PHQ_Clinical_2018'))

setnames(Outcomes_Restricted_2019, old = c('Insurance', 'FPL_Level', 'On_Hormones', 'On_PrEP',  
                                'HIV_Status', 'HIV_Suppressed', 'PHQ_Clinical'), 
         new = c('Insurance_2019', 'FPL_Level_2019', 'On_Hormones_2019', 'On_PrEP_2019', 
                                'HIV_Status_2019', 'HIV_Suppressed_2019', 'PHQ_Clinical_2019'))

```

```{r}
#Left joining the datasets on patient ID

Demo_16_Restricted <- merge(Cohort_Demographics_Restricted, Outcomes_Restricted_2016, by = "ID")

Demo_1617_Restricted <- merge(Demo_16_Restricted, Outcomes_Restricted_2017, by = "ID", all.x = TRUE)

Demo_161718_Restricted <- merge(Demo_1617_Restricted, Outcomes_Restricted_2018, by = "ID", all.x = TRUE)

general_cohort_restricted_wide <- merge(Demo_161718_Restricted, Outcomes_Restricted_2019, by = "ID", all.x = TRUE)
```

```{r}
#Converting character variables to factor variables
general_cohort_restricted_wide$Gender <- factor(general_cohort_restricted_wide$Gender)
general_cohort_restricted_wide$Race <- factor(general_cohort_restricted_wide$Race)
general_cohort_restricted_wide$Ethnicity <- factor(general_cohort_restricted_wide$Ethnicity)

general_cohort_restricted_wide$Insurance_2016 <- factor(general_cohort_restricted_wide$Insurance_2016)
general_cohort_restricted_wide$FPL_Level_2016 <- factor(general_cohort_restricted_wide$FPL_Level_2016)
general_cohort_restricted_wide$HIV_Suppressed_2016 <- factor(general_cohort_restricted_wide$HIV_Suppressed_2016)
general_cohort_restricted_wide$PHQ_Clinical_2016 <- factor(general_cohort_restricted_wide$PHQ_Clinical_2016)

general_cohort_restricted_wide$Insurance_2017 <- factor(general_cohort_restricted_wide$Insurance_2017)
general_cohort_restricted_wide$FPL_Level_2017 <- factor(general_cohort_restricted_wide$FPL_Level_2017)
general_cohort_restricted_wide$HIV_Suppressed_2017 <- factor(general_cohort_restricted_wide$HIV_Suppressed_2017)
general_cohort_restricted_wide$PHQ_Clinical_2017 <- factor(general_cohort_restricted_wide$PHQ_Clinical_2017)

general_cohort_restricted_wide$Insurance_2018 <- factor(general_cohort_restricted_wide$Insurance_2018)
general_cohort_restricted_wide$FPL_Level_2018 <- factor(general_cohort_restricted_wide$FPL_Level_2018)
general_cohort_restricted_wide$HIV_Suppressed_2018 <- factor(general_cohort_restricted_wide$HIV_Suppressed_2018)
general_cohort_restricted_wide$PHQ_Clinical_2018 <- factor(general_cohort_restricted_wide$PHQ_Clinical_2018)

general_cohort_restricted_wide$Insurance_2019 <- factor(general_cohort_restricted_wide$Insurance_2019)
general_cohort_restricted_wide$FPL_Level_2019 <- factor(general_cohort_restricted_wide$FPL_Level_2019)
general_cohort_restricted_wide$HIV_Suppressed_2019 <- factor(general_cohort_restricted_wide$HIV_Suppressed_2019)
general_cohort_restricted_wide$PHQ_Clinical_2019 <- factor(general_cohort_restricted_wide$PHQ_Clinical_2019)
```

```{r}
median(general_cohort_restricted_wide$Age, na.rm = TRUE)
IQR(general_cohort_restricted_wide$Age, na.rm = TRUE)
```

## Examining Missingness in Restricted (Non-Imputed) Dataset

```{r}
miss_check_data <- general_cohort_restricted_wide[c('ID', "Site", "Age", "Gender", "Race", "Ethnicity", "cohort_years", "Insurance_2016", "FPL_Level_2016", "On_Hormones_2016", "On_PrEP_2016", "HIV_Status_2016", "HIV_Suppressed_2016", "PHQ_Clinical_2016")]
```

```{r}
miss_check_data$miss_n <- rowSums(is.na(miss_check_data))

miss_check_data$any_miss <- ifelse(miss_check_data$miss_n >= 1, 1, 0)

table(miss_check_data$any_miss)
```

### Multiple Imputation with Chained Equations

```{r}
#Establishing imputation methods and an initial predictor matrix for the final imputation model

library(mice)

init = mice(general_cohort_restricted_wide, maxit=0, seed=0829) 
meth = init$method

meth[c("ID")]=""
meth[c("Site")]=""
meth[c("Age")]="pmm"
meth[c("Gender")]="polyreg"
meth[c("Race")]="polyreg"
meth[c("Ethnicity")]="cart"
meth[c("cohort_years")]=""

meth[c("Insurance_2016")]="polyreg"
meth[c("FPL_Level_2016")]="cart"
meth[c("On_Hormones_2016")]="cart"
meth[c("On_PrEP_2016")]="cart"
meth[c("HIV_Status_2016")]="cart"
meth[c("HIV_Suppressed_2016")]="cart"
meth[c("PHQ_Clinical_2016")]=""

meth[c("Insurance_2017")]="polyreg"
meth[c("FPL_Level_2017")]="cart"
meth[c("On_Hormones_2017")]="cart"
meth[c("On_PrEP_2017")]="cart"
meth[c("HIV_Status_2017")]="cart"
meth[c("HIV_Suppressed_2017")]="cart"
meth[c("PHQ_Clinical_2017")]=""

meth[c("Insurance_2018")]="polyreg"
meth[c("FPL_Level_2018")]="cart"
meth[c("On_Hormones_2018")]="cart"
meth[c("On_PrEP_2018")]="cart"
meth[c("HIV_Status_2018")]="cart"
meth[c("HIV_Suppressed_2018")]="cart"
meth[c("PHQ_Clinical_2018")]=""

meth[c("Insurance_2019")]="polyreg"
meth[c("FPL_Level_2019")]="cart"
meth[c("On_Hormones_2019")]="cart"
meth[c("On_PrEP_2019")]="cart"
meth[c("HIV_Status_2019")]="cart"
meth[c("HIV_Suppressed_2019")]="cart"
meth[c("PHQ_Clinical_2019")]=""

predM_Restricted = init$predictorMatrix
```

```{r}
#Editing the predictor matrix to remove IDs and post-baseline Outcomes_Restricted from the imputation model
predM_Restricted[,1] <- 0
predM_Restricted[,15:35] <- 0
```

```{r, echo = FALSE, eval = FALSE}
#Applying multiple imputation with chained equations to the wide cohort dataset.
library(mice)

imputed_data_restricted <- mice(general_cohort_restricted_wide, method=meth, predictorMatrix = predM_Restricted, m=5, seed = 0829)
```

```{r}
wide_imputed_dataset_restricted <- complete(imputed_data_restricted, action="long", include=FALSE)
```

```{r}
library(miceadds)

#Checking missingness in the imputed datasets
data.frame(prop_miss(wide_imputed_dataset_restricted))
```

### Cleaning Imputed dataset_restricted

```{r}
#Examining raw cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset_restricted$HIV_Status_2016, wide_imputed_dataset_restricted$On_PrEP_2016)
table(wide_imputed_dataset_restricted$HIV_Status_2017, wide_imputed_dataset_restricted$On_PrEP_2017)
table(wide_imputed_dataset_restricted$HIV_Status_2018, wide_imputed_dataset_restricted$On_PrEP_2018)
table(wide_imputed_dataset_restricted$HIV_Status_2019, wide_imputed_dataset_restricted$On_PrEP_2019)
```

```{r}
#Recoding on PrEP variable to "No" for folks who are HIV+
wide_imputed_dataset_restricted$On_PrEP_2016[wide_imputed_dataset_restricted$HIV_Status_2016 == "HIV+"] <- "No"
wide_imputed_dataset_restricted$On_PrEP_2017[wide_imputed_dataset_restricted$HIV_Status_2017 == "HIV+"] <- "No"
wide_imputed_dataset_restricted$On_PrEP_2018[wide_imputed_dataset_restricted$HIV_Status_2018 == "HIV+"] <- "No"
wide_imputed_dataset_restricted$On_PrEP_2019[wide_imputed_dataset_restricted$HIV_Status_2019 == "HIV+"] <- "No"
```

```{r}
#Examining cleaned cross-tabs for HIV status and being on PrEP
table(wide_imputed_dataset_restricted$HIV_Status_2016, wide_imputed_dataset_restricted$On_PrEP_2016)
table(wide_imputed_dataset_restricted$HIV_Status_2017, wide_imputed_dataset_restricted$On_PrEP_2017)
table(wide_imputed_dataset_restricted$HIV_Status_2018, wide_imputed_dataset_restricted$On_PrEP_2018)
table(wide_imputed_dataset_restricted$HIV_Status_2019, wide_imputed_dataset_restricted$On_PrEP_2019)
```

```{r}
#Examining raw cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset_restricted$HIV_Status_2016, wide_imputed_dataset_restricted$HIV_Suppressed_2016)
table(wide_imputed_dataset_restricted$HIV_Status_2017, wide_imputed_dataset_restricted$HIV_Suppressed_2017)
table(wide_imputed_dataset_restricted$HIV_Status_2018, wide_imputed_dataset_restricted$HIV_Suppressed_2018)
table(wide_imputed_dataset_restricted$HIV_Status_2019, wide_imputed_dataset_restricted$HIV_Suppressed_2019)
```

```{r}
#Recoding viral suppression variable to "<200 copies" for folks who are HIV-
wide_imputed_dataset_restricted$HIV_Suppressed_2016[wide_imputed_dataset_restricted$HIV_Status_2016 == "HIV-"] <- "<200 copies"
wide_imputed_dataset_restricted$HIV_Suppressed_2017[wide_imputed_dataset_restricted$HIV_Status_2017 == "HIV-"] <- "<200 copies"
wide_imputed_dataset_restricted$HIV_Suppressed_2018[wide_imputed_dataset_restricted$HIV_Status_2018 == "HIV-"] <- "<200 copies"
wide_imputed_dataset_restricted$HIV_Suppressed_2019[wide_imputed_dataset_restricted$HIV_Status_2019 == "HIV-"] <- "<200 copies"
```

```{r}
#Examining cleaned cross-tabs for HIV status and viral suppression
table(wide_imputed_dataset_restricted$HIV_Status_2016, wide_imputed_dataset_restricted$HIV_Suppressed_2016)
table(wide_imputed_dataset_restricted$HIV_Status_2017, wide_imputed_dataset_restricted$HIV_Suppressed_2017)
table(wide_imputed_dataset_restricted$HIV_Status_2018, wide_imputed_dataset_restricted$HIV_Suppressed_2018)
table(wide_imputed_dataset_restricted$HIV_Status_2019, wide_imputed_dataset_restricted$HIV_Suppressed_2019)
```

```{r}
wide_imputed_dataset_restricted <- mutate(wide_imputed_dataset_restricted, HIV_Risk_2016 = case_when(
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "No") ~ 0,
  (HIV_Status_2016 == "HIV-") & (On_PrEP_2016 == "Yes") ~ 1,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "<200 copies") ~ 2,
  (HIV_Status_2016 == "HIV+") & (HIV_Suppressed_2016 == "200+ copies") ~ 3
))

wide_imputed_dataset_restricted$HIV_Risk_2016 <- factor(wide_imputed_dataset_restricted$HIV_Risk_2016)

levels(wide_imputed_dataset_restricted$HIV_Risk_2016) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset_restricted <- mutate(wide_imputed_dataset_restricted, HIV_Risk_2017 = case_when(
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "No") ~ 0,
  (HIV_Status_2017 == "HIV-") & (On_PrEP_2017 == "Yes") ~ 1,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "<200 copies") ~ 2,
  (HIV_Status_2017 == "HIV+") & (HIV_Suppressed_2017 == "200+ copies") ~ 3
))

wide_imputed_dataset_restricted$HIV_Risk_2017 <- factor(wide_imputed_dataset_restricted$HIV_Risk_2017)

levels(wide_imputed_dataset_restricted$HIV_Risk_2017) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset_restricted <- mutate(wide_imputed_dataset_restricted, HIV_Risk_2018 = case_when(
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "No") ~ 0,
  (HIV_Status_2018 == "HIV-") & (On_PrEP_2018 == "Yes") ~ 1,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "<200 copies") ~ 2,
  (HIV_Status_2018 == "HIV+") & (HIV_Suppressed_2018 == "200+ copies") ~ 3
))

wide_imputed_dataset_restricted$HIV_Risk_2018 <- factor(wide_imputed_dataset_restricted$HIV_Risk_2018)

levels(wide_imputed_dataset_restricted$HIV_Risk_2018) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
wide_imputed_dataset_restricted <- mutate(wide_imputed_dataset_restricted, HIV_Risk_2019 = case_when(
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "No") ~ 0,
  (HIV_Status_2019 == "HIV-") & (On_PrEP_2019 == "Yes") ~ 1,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "<200 copies") ~ 2,
  (HIV_Status_2019 == "HIV+") & (HIV_Suppressed_2019 == "200+ copies") ~ 3
))

wide_imputed_dataset_restricted$HIV_Risk_2019 <- factor(wide_imputed_dataset_restricted$HIV_Risk_2019)

levels(wide_imputed_dataset_restricted$HIV_Risk_2019) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")
```

```{r}
table(wide_imputed_dataset_restricted$HIV_Risk_2016)
table(wide_imputed_dataset_restricted$HIV_Risk_2017)
table(wide_imputed_dataset_restricted$HIV_Risk_2018)
table(wide_imputed_dataset_restricted$HIV_Risk_2019)
```

## Creating Imputed Repeated Measures dataset_restricted

```{r}
#Creating placeholder variables for imputation and ID keys
wide_imputed_dataset_restricted$Imp_Num <- wide_imputed_dataset_restricted$".imp"
wide_imputed_dataset_restricted$Imp_ID <- wide_imputed_dataset_restricted$".id"
```

```{r}
impute_data_restricted_2016 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2016 as Insurance, FPL_Level_2016 as FPL_Level, On_Hormones_2016 as On_Hormones, On_PrEP_2016 as On_PrEP, HIV_Status_2016 as HIV_Status,
HIV_Suppressed_2016 as HIV_Suppressed, PHQ_Clinical_2016 as PHQ_Clinical FROM wide_imputed_dataset_restricted ORDER BY Imp_Num")

impute_data_restricted_2016$Year <- 2016
```

```{r}
impute_data_restricted_2017 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2017 as Insurance, FPL_Level_2017 as FPL_Level, On_Hormones_2017 as On_Hormones, On_PrEP_2017 as On_PrEP, HIV_Status_2017 as HIV_Status,
HIV_Suppressed_2017 as HIV_Suppressed, PHQ_Clinical_2017 as PHQ_Clinical FROM wide_imputed_dataset_restricted ORDER BY Imp_Num")

impute_data_restricted_2017$Year <- 2017
```

```{r}
#Creating placeholder variables for imputation and ID keys
wide_imputed_dataset_restricted$Imp_Num <- wide_imputed_dataset_restricted$".imp"
wide_imputed_dataset_restricted$Imp_ID <- wide_imputed_dataset_restricted$".id"
```

```{r}
impute_data_restricted_2018 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2018 as Insurance, FPL_Level_2018 as FPL_Level, On_Hormones_2018 as On_Hormones, On_PrEP_2018 as On_PrEP, HIV_Status_2018 as HIV_Status,
HIV_Suppressed_2018 as HIV_Suppressed, PHQ_Clinical_2018 as PHQ_Clinical FROM wide_imputed_dataset_restricted ORDER BY Imp_Num")

impute_data_restricted_2018$Year <- 2018
```

```{r}
impute_data_restricted_2019 <- sqldf("SELECT Imp_Num, Imp_ID, ID, Site, Age, Gender, Race, Ethnicity, cohort_years, Insurance_2019 as Insurance, FPL_Level_2019 as FPL_Level, On_Hormones_2019 as On_Hormones, On_PrEP_2019 as On_PrEP, HIV_Status_2019 as HIV_Status,
HIV_Suppressed_2019 as HIV_Suppressed, PHQ_Clinical_2019 as PHQ_Clinical FROM wide_imputed_dataset_restricted ORDER BY Imp_Num")

impute_data_restricted_2019$Year <- 2019
```

```{r}
long_imputed_data_restricted <- rbind(impute_data_restricted_2016, impute_data_restricted_2017, impute_data_restricted_2018, impute_data_restricted_2019)
```

```{r}
#Adding a variable for observation year
long_imputed_data_restricted <- mutate(long_imputed_data_restricted, Obs_Year = case_when(
  Year == 2016 ~ 0,
  Year == 2017 ~ 1,
  Year == 2018 ~ 2,
  Year == 2019 ~ 3
))

#Creating a categorical variable for age
long_imputed_data_restricted <- mutate(long_imputed_data_restricted, AgeCat = case_when(
  (18 <= Age) & (Age <= 24) ~ 0,
  (25 <= Age) & (Age <= 30) ~ 1,
  (31 <= Age) & (Age <= 40) ~ 2,
  (41 <= Age) & (Age <= 50) ~ 3,
  (Age >= 51) ~ 4))

long_imputed_data_restricted$AgeCat <- factor(long_imputed_data_restricted$AgeCat, levels = c(0, 1, 2, 3, 4), 
                                 labels = c("18-24 years", "25-30 years", "31-40 years",
                                            "41-50 years", "51+ years"))
```

```{r}
table(long_imputed_data_restricted$Race)
```


```{r}
#Collapsing race categories

long_imputed_data_restricted <- mutate(long_imputed_data_restricted, Race_Three = case_when(
  Race == "White" ~ 0,
  Race == "Asian" ~ 1,
  Race == "Black or African American" ~ 2,
  Race == "American Indian or Alaska Native" | Race == "Another Race" | Race == "Multiracial" | Race == "Native Hawaiian or Other Pacific Islander" ~ 3
))

long_imputed_data_restricted$Race_Three <- factor(long_imputed_data_restricted$Race_Three)
levels(long_imputed_data_restricted$Race_Three) <- c("White", "Asian", "Black or African American", "Another Race")

table(long_imputed_data_restricted$Race_Three)
```


```{r}
#Recoding imputation and ID keys
long_imputed_data_restricted$".imp" <- long_imputed_data_restricted$Imp_Num
long_imputed_data_restricted$".id" <- long_imputed_data_restricted$Imp_ID

#Dropping placeholder keys
long_imputed_data_restricted <- subset(long_imputed_data_restricted, select = -c(Imp_Num, Imp_ID))
```

```{r}
long_imputed_data_restricted <- long_imputed_data_restricted[order(long_imputed_data_restricted$.imp, long_imputed_data_restricted$.id),]
```

## Cleaning Long Imputed dataset_restricted

```{r}
#Recoding character variables as factors
long_imputed_data_restricted$Insurance <- factor(long_imputed_data_restricted$Insurance)
levels(long_imputed_data_restricted$Insurance)

long_imputed_data_restricted$FPL_Level <- factor(long_imputed_data_restricted$FPL_Level)
levels(long_imputed_data_restricted$FPL_Level)

long_imputed_data_restricted$On_Hormones <- factor(long_imputed_data_restricted$On_Hormones)
levels(long_imputed_data_restricted$On_Hormones)

long_imputed_data_restricted$On_PrEP <- factor(long_imputed_data_restricted$On_PrEP)
levels(long_imputed_data_restricted$On_PrEP)

long_imputed_data_restricted$HIV_Status <- factor(long_imputed_data_restricted$HIV_Status)
levels(long_imputed_data_restricted$HIV_Status)

long_imputed_data_restricted$HIV_Suppressed <- factor(long_imputed_data_restricted$HIV_Suppressed)
levels(long_imputed_data_restricted$HIV_Suppressed)

long_imputed_data_restricted$PHQ_Clinical <- factor(long_imputed_data_restricted$PHQ_Clinical)
levels(long_imputed_data_restricted$PHQ_Clinical)
```

```{r}
long_imputed_data_restricted <- mutate(long_imputed_data_restricted, HIV_Risk = case_when(
  (HIV_Status == "HIV-") & (On_PrEP == "No") ~ 0,
  (HIV_Status == "HIV-") & (On_PrEP == "Yes") ~ 1,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "<200 copies") ~ 2,
  (HIV_Status == "HIV+") & (HIV_Suppressed == "200+ copies") ~ 3
))

long_imputed_data_restricted$HIV_Risk <- factor(long_imputed_data_restricted$HIV_Risk)

levels(long_imputed_data_restricted$HIV_Risk) <- c("HIV- & Not on PrEP", "HIV- & On PrEP", 
                                     "HIV+ & Virally Suppressed", 
                                     "HIV+ & Non-Virally Suppressed")

```

```{r}
library(dplyr)

long_imputed_data_restricted <- mutate(long_imputed_data_restricted, SAB = case_when(
  Gender == "Male" | Gender == "Nonbinary AFAB" ~ 0,
  Gender == "Female" | Gender == "Nonbinary AMAB" ~ 1
))

long_imputed_data_restricted$SAB <- factor(long_imputed_data_restricted$SAB)
levels(long_imputed_data_restricted$SAB) <- c("AFAB", "AMAB")

table(long_imputed_data_restricted$SAB, useNA = 'always')
```


```{r}
long_imputed_data_restricted <- mutate(long_imputed_data_restricted, PHQ_Numeric = case_when(
  PHQ_Clinical == "No" ~ 0,
  PHQ_Clinical == "Yes" ~ 1
))
```

## Creating Site Stratified Datasets

```{r}
fh_long_data <- long_imputed_data_restricted[which(long_imputed_data_restricted$Site == "FH"),]

cl_long_data <- long_imputed_data_restricted[which(long_imputed_data_restricted$Site == "CL"),]
```


## Creating dataset_restricted List

```{r}
imputed_data_restricted_list <- split(long_imputed_data_restricted, f = long_imputed_data_restricted$.imp)

fh_long_data_list <- split(fh_long_data, f = fh_long_data$.imp)

cl_long_data_list <- split(cl_long_data, f = cl_long_data$.imp)
```

## GEE Modeling Across Datasets

```{r}
library(miceadds)

imputed_mids_restricted <- datalist2mids(imputed_data_restricted_list)
```

## Bivariate Models (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_hormones_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hormones_restricted <- pool(PHQ_hormones_restricted)

options(digits = 3)

summary(PHQ_combined_hormones_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_site_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ Site + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_site_restricted <- pool(PHQ_site_restricted)

options(digits = 3)

summary(PHQ_combined_site_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_age_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_age_restricted <- pool(PHQ_age_restricted)

options(digits = 3)

summary(PHQ_combined_age_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_gender_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ relevel(Gender, ref = "Male") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_gender_restricted <- pool(PHQ_gender_restricted)

options(digits = 3)

summary(PHQ_combined_gender_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_race_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ relevel(Race, ref = "White") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_race_restricted <- pool(PHQ_race_restricted)

options(digits = 3)

summary(PHQ_combined_race_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_ethnicity_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ relevel(Ethnicity, ref = "Not Hispanic") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_ethnicity_restricted <- pool(PHQ_ethnicity_restricted)

options(digits = 3)

summary(PHQ_combined_ethnicity_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_insurance_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ Insurance + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_insurance_restricted <- pool(PHQ_insurance_restricted)

options(digits = 3)

summary(PHQ_combined_insurance_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_fpl_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ relevel(FPL_Level, ref = "251%+") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_fpl_restricted <- pool(PHQ_fpl_restricted)

options(digits = 3)

summary(PHQ_combined_fpl_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_hiv_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ HIV_Risk + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hiv_restricted <- pool(PHQ_hiv_restricted)

options(digits = 3)

summary(PHQ_combined_hiv_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_years_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_years_restricted <- pool(PHQ_years_restricted)

options(digits = 3)

summary(PHQ_combined_years_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

## Multivariable Model (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_adjusted_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Site + AgeCat + relevel(Gender, ref = "Male") + relevel(Race, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+") + 
                                                     HIV_Risk + cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_adjusted_restricted <- pool(PHQ_adjusted_restricted)

options(digits = 3)

summary(PHQ_combined_adjusted_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_adjusted_restricted_2 <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Site + AgeCat + relevel(SAB, ref = "AFAB") + relevel(Race, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+") + 
                                                     HIV_Risk + cohort_years + Obs_Year + On_Hormones*relevel(SAB, ref = "AFAB"), family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_adjusted_restricted_2 <- pool(PHQ_adjusted_restricted_2)

options(digits = 3)
options(scipen = 9)

summary(PHQ_combined_adjusted_restricted_2, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(EValue)

evalues.RR(0.838, lo = 0.712, hi = 0.985)
```


## Interaction Models

```{r}
library(mice)
library(geepack)

PHQ_ageint_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + AgeCat + On_Hormones*AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_ageint_restricted <- pool(PHQ_ageint_restricted)

options(digits = 3)

summary(PHQ_combined_ageint_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```


```{r}
library(mice)
library(geepack)

PHQ_genderint_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + relevel(Gender, ref = "Male") + On_Hormones*relevel(Gender, ref = "Male") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_genderint_restricted <- pool(PHQ_genderint_restricted)

options(digits = 3)

summary(PHQ_combined_genderint_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_sexint_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + relevel(SAB, ref = "AFAB") + On_Hormones*relevel(SAB, ref = "AFAB") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_sexint_restricted <- pool(PHQ_sexint_restricted)

options(digits = 3)

summary(PHQ_combined_sexint_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_raceint_restricted <- with(imputed_mids_restricted, exp <- geeglm(PHQ_Numeric ~ On_Hormones + relevel(Race_Three, ref = "White") + On_Hormones*relevel(Race_Three, ref = "White") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_raceint_restricted <- pool(PHQ_raceint_restricted)

options(digits = 3)

summary(PHQ_combined_raceint_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```


## Fenway Health Models

```{r}
library(miceadds)

imputed_mids_fh <- datalist2mids(fh_long_data_list)
```

## Bivariate Models (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_hormones_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hormones_restricted <- pool(PHQ_hormones_restricted)

options(digits = 3)

summary(PHQ_combined_hormones_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_age_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_age_restricted <- pool(PHQ_age_restricted)

options(digits = 3)

summary(PHQ_combined_age_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_gender_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ relevel(Gender, ref = "Male") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_gender_restricted <- pool(PHQ_gender_restricted)

options(digits = 3)

summary(PHQ_combined_gender_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_race_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ relevel(Race, ref = "White") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_race_restricted <- pool(PHQ_race_restricted)

options(digits = 3)

summary(PHQ_combined_race_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_ethnicity_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ relevel(Ethnicity, ref = "Not Hispanic") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_ethnicity_restricted <- pool(PHQ_ethnicity_restricted)

options(digits = 3)

summary(PHQ_combined_ethnicity_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_insurance_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ Insurance + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_insurance_restricted <- pool(PHQ_insurance_restricted)

options(digits = 3)

summary(PHQ_combined_insurance_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_fpl_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ relevel(FPL_Level, ref = "251%+") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_fpl_restricted <- pool(PHQ_fpl_restricted)

options(digits = 3)

summary(PHQ_combined_fpl_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_hiv_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ HIV_Risk + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hiv_restricted <- pool(PHQ_hiv_restricted)

options(digits = 3)

summary(PHQ_combined_hiv_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_years_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_years_restricted <- pool(PHQ_years_restricted)

options(digits = 3)

summary(PHQ_combined_years_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

## Multivariable Model (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_adjusted_restricted <- with(imputed_mids_fh, exp <- geeglm(PHQ_Numeric ~ On_Hormones + AgeCat + relevel(Gender, ref = "Male") + relevel(Race, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+") + 
                                                     HIV_Risk + cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_adjusted_restricted <- pool(PHQ_adjusted_restricted)

options(digits = 3)

summary(PHQ_combined_adjusted_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```


## Callen-Lorde Models

```{r}
library(miceadds)

imputed_mids_cl <- datalist2mids(cl_long_data_list)
```

## Bivariate Models (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_hormones_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ On_Hormones + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hormones_restricted <- pool(PHQ_hormones_restricted)

options(digits = 3)

summary(PHQ_combined_hormones_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_age_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ AgeCat + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_age_restricted <- pool(PHQ_age_restricted)

options(digits = 3)

summary(PHQ_combined_age_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_gender_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ relevel(Gender, ref = "Male") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_gender_restricted <- pool(PHQ_gender_restricted)

options(digits = 3)

summary(PHQ_combined_gender_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_race_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ relevel(Race_Three, ref = "White") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_race_restricted <- pool(PHQ_race_restricted)

options(digits = 3)

summary(PHQ_combined_race_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_ethnicity_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ relevel(Ethnicity, ref = "Not Hispanic") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_ethnicity_restricted <- pool(PHQ_ethnicity_restricted)

options(digits = 3)

summary(PHQ_combined_ethnicity_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_insurance_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ Insurance + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_insurance_restricted <- pool(PHQ_insurance_restricted)

options(digits = 3)

summary(PHQ_combined_insurance_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_fpl_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ relevel(FPL_Level, ref = "251%+") + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_fpl_restricted <- pool(PHQ_fpl_restricted)

options(digits = 3)

summary(PHQ_combined_fpl_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_hiv_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ HIV_Risk + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_hiv_restricted <- pool(PHQ_hiv_restricted)

options(digits = 3)

summary(PHQ_combined_hiv_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

```{r}
library(mice)
library(geepack)

PHQ_years_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ cohort_years + Obs_Year, family = poisson(link = "log"), 
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_years_restricted <- pool(PHQ_years_restricted)

options(digits = 3)

summary(PHQ_combined_years_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```

## Multivariable Model (Restricted)

```{r}
library(mice)
library(geepack)

PHQ_adjusted_restricted <- with(imputed_mids_cl, exp <- geeglm(PHQ_Numeric ~ On_Hormones + AgeCat + relevel(Gender, ref = "Male") + relevel(Race_Three, ref = "White") + 
                                                   relevel(Ethnicity, ref = "Not Hispanic") + Insurance + relevel(FPL_Level, ref = "251%+") + 
                                                     HIV_Risk + cohort_years + Obs_Year, family = poisson(link = "log"),
                                                 id = ID, waves = Obs_Year, corstr = "exchangeable"))
```

```{r}
#Pooling estimates across datasets

PHQ_combined_adjusted_restricted <- pool(PHQ_adjusted_restricted)

options(digits = 3)

summary(PHQ_combined_adjusted_restricted, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
```